%\documentclass[mathserif]{beamer}
\documentclass[handout]{beamer}
%\usetheme{Goettingen}
\usetheme{Warsaw}
%\usetheme{Singapore}
%\usetheme{Frankfurt}
%\usetheme{Copenhagen}
%\usetheme{Szeged}
%\usetheme{Montpellier}
%\usetheme{CambridgeUS}
%\usecolortheme{}
%\setbeamercovered{transparent}
\usepackage[english, activeacute]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath, amssymb}
\usepackage{dsfont}
\usepackage{graphics}
\usepackage{cases}
\usepackage{graphicx}
\usepackage{pgf}
\usepackage{epsfig}
\usepackage{amssymb}
\usepackage{multirow}	
\usepackage{amstext}
\usepackage[ruled,vlined,lined]{algorithm2e}
\usepackage{amsmath}
\usepackage{epic}
\usepackage{epsfig}
\usepackage{fontenc}
\usepackage{framed,color}
\usepackage{palatino, url, multicol}
\usepackage{listings}
%\algsetup{indent=2em}


\vspace{-0.5cm}
\title{Model Evaluation and Information Criteria}
\vspace{-0.5cm}
\author[Felipe Bravo Márquez]{\footnotesize
%\author{\footnotesize  
 \textcolor[rgb]{0.00,0.00,1.00}{Felipe José Bravo Márquez}} 
\date{ \today }




\begin{document}
\begin{frame}
\titlepage


\end{frame}


%%%%%%%%%%%%%%%%%%%%%%%%%%%


\begin{frame}{Model Evaluation and Information Criteria}
\scriptsize{
\begin{itemize}
\item In the context of scientific models, there are two fundamental kinds of statistical error \cite{mcelreath2020statistical}:

\begin{itemize}\scriptsize{
 \item \textbf{Overfitting}, which leads to poor prediction by learning too much from the data.
\item \textbf{Underfitting}, which leads to poor prediction by learning too little from the data.}
\end{itemize}

\item There are two common families of approaches to tackle these problems.


\begin{itemize}\scriptsize{
 \item \textbf{Regularization}: a mechanism to tell our models not to get too excited by the data.
\item \textbf{Information criteria}: a scoring device to estimate predictive accuracy of our models.}
\end{itemize}

\item In order to introduce information criteria, this class must also introduce \textbf{information theory}.

 
\end{itemize}



} 

\end{frame}



\begin{frame}{The problem with parameters}
\scriptsize{

\begin{itemize}
\item In the class of linear regression we learned that including more attributes can lead to a more accurate model.

\item However, we have also learned that adding more variables almost always improves the fit of the model to the data, as measured by the coefficient of determination $R^2$. 
\item This is true even when the variables you add to a model
are just random numbers, with no relation to the outcome. 

\item So it's no good to choose among models using only fit to the data.






\end{itemize}


} 
\end{frame}

\begin{frame}{The problem with parameters}
\scriptsize{

\begin{itemize}

\item While more complex models fit the data better, they often predict new data worse.

\item This means that a complex model will be very sensitive to the exact sample used to fit it.

\item This will lead to potentially large mistakes when future data is not exactly like the past data.

\item But simple models, with too few parameters, tend instead to underfit, systematically over-predicting or under-predicting the data.

\item Regardless of how well future data resemble past data. 
\item So we can't always favor either simple models or complex models.

\item Let’s examine both of these issues in the context of a simple data example.

\end{itemize}


} 
\end{frame}


\begin{frame}[fragile]{The problem with parameters}
\scriptsize{

\begin{itemize}

\item We are going to create a data.frame containing  average brain volumes and body masses for seven hominin species.

\begin{verbatim}
sppnames <- c( "afarensis","africanus","habilis",
               "boisei", "rudolfensis","ergaster",
               "sapiens")
brainvolcc <- c( 438 , 452 , 612, 521, 752, 871, 
                 1350 )
masskg <- c( 37.0 , 35.5 , 34.5 , 41.5 , 55.5 , 
             61.0 , 53.5 )
d <- data.frame( species=sppnames , brain=brainvolcc,
                 mass=masskg ) 
\end{verbatim}

\item It's not unusual for data like this to be highly correlated.
\item Brain size is correlated with body size, across
species. 
\item We will model brain size as a linear function of body size.

\item We will fit a series of increasingly complex model families and see which function fits the data best.

\end{itemize}


} 
\end{frame}


\begin{frame}{Information theory and model performance}
\scriptsize{

\begin{itemize}
\item Blabla
\end{itemize}


} 
\end{frame}


\begin{frame}{Regularization}
\scriptsize{

\begin{itemize}
\item Blabla
\end{itemize}


} 
\end{frame}

\begin{frame}{Information criteria}
\scriptsize{

\begin{itemize}
\item Blabla
\end{itemize}


} 
\end{frame}


\begin{frame}{Using information criteria}
\scriptsize{

\begin{itemize}
\item Blabla
\end{itemize}


} 
\end{frame}




\begin{frame}{Conclusions}
\scriptsize{

\begin{itemize}
\item Blabla
\end{itemize}


} 
\end{frame}


%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[allowframebreaks]\scriptsize
\frametitle{References}
\bibliography{bio}
\bibliographystyle{apalike}
%\bibliographystyle{flexbib}
\end{frame}  









%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}
